Modal type theory based on the intuitionistic epistemic logic

### 2018-01-13 article.tex
added some fixes and rewrite

@@

  \begin{center}
  \begin{prooftree}
  \AxiomC{$ $}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\Gamma , x : \alpha \vdash x : \alpha$}
  \end{prooftree}
  \end{center}

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  \begin{prooftree}
  \AxiomC{$\Gamma, x : \alpha \vdash M : \beta$}
  \RightLabel{$\rightarrow_i$}
  \UnaryInfC{$\Gamma \vdash \lambda x. M : \alpha \to \beta$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{ $\Gamma \vdash x : \alpha$ }
  \AxiomC{ $\Gamma \vdash y : \beta$ }
  \RightLabel{$\times_i$}
  \BinaryInfC{$\Gamma \vdash (x, y) : \alpha \times \beta$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash x : \alpha$}
    \RightLabel{$\textbf{K}_I$}
    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}\alpha $}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}

  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
  \begin{prooftree}
  \AxiomC{$\Gamma \vdash f : \alpha \to \beta$}
  \AxiomC{$\Gamma \vdash x : \alpha$}
  \RightLabel{$\rightarrow_e$}
  \BinaryInfC{$\Gamma \vdash fx : \beta$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{ $\Gamma \vdash p : \alpha_1 \times \alpha_2$ }
  \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
  \UnaryInfC{$\Gamma \vdash \pi_i p : \alpha_i$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
    \AxiomC{$\vec{N} : \vec{A} \vdash M : B$}
    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M : {\bf K} B$}
  \end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

  7) $FV({\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
  \end{defin}

@@

  In $\lambda \textbf{K}$ we have the following computational rules. We will define them for terms with
  ${\bf pure \: }$ or $\star$. Reduction rules for $(,)$ and $\pi_i$ are described, for example, in
  \cite{Pierce}.

@@

@@

  Let $\Gamma = \{ x_1 : \alpha_1, ..., x_n : \alpha_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.

@@

  If $\Gamma \vdash M : \alpha$, then $FV(M) \subseteq dom(\Gamma)$

@@

  Induction on the derivation of $\Gamma \vdash M : \alpha$.

@@

  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ implies that $\Gamma \vdash M : \alpha$;

@@

  Let $\Gamma \vdash M : \alpha$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : \alpha$.

@@

  1) Let $\Gamma \vdash x : \alpha$, such that $\Gamma = \Delta, x : \alpha$ and $\Theta \subseteq \Gamma$.
  Let $\Sigma = \Theta \setminus \Gamma$, or, which is the same, $\Sigma = \Theta \setminus \Delta, x :
  \alpha$, then $\Sigma,
  \Delta, x : \alpha \vdash x : \alpha$, or, $\Theta \vdash x : \alpha$.

  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \subseteq \Theta$.

  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$, then $\Gamma \vdash M : \alpha$ by generation and, by
  hypothesis, $\Theta \vdash M : \alpha$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}\alpha$ by applying
  $\textbf{K}_I$-rule.

  3) Let $\Gamma \vdash M \star N : \textbf{K}\beta$. So $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
  \vdash N : \textbf{K}\alpha$. By hypothesis $\Delta \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Delta \vdash N :
  \textbf{K}\alpha$. Then $\Delta \vdash M \star N : \textbf{K}\beta$.

@@

  If $\Gamma \vdash M : \alpha$, then $\Gamma \uparrow FV(M) \vdash M : \alpha$, where $\Gamma \uparrow
  FV(M)$ is a subcontext of $\Gamma$, such that $dom(\Gamma \uparrow FV(M)) = dom(\Gamma) \cap FV(M)$.

@@

  Induction by derivation. We consider the base of induction and the case with $\textbf{K}_{app}$. The rest
  cases are
  proven by the same way.

@@

  1) Let $\Gamma \vdash x : \alpha$, where $\Gamma = \Delta, x : \alpha$, $x \in \mathbb{V}$.

  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : \alpha) \uparrow FV(x) = \{ x
  : \alpha\}$, then $x : \alpha \vdash x : \alpha$ by axiom.

  2) Let $\Gamma \vdash M \star N : \textbf{K}\beta$.

  By generation $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash N : \textbf{K}\alpha$.

  By induction hypothesis $\Gamma \uparrow FV(M) \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
  \uparrow FV(N) \vdash N : \textbf{K}\beta$.

  But $\Gamma \uparrow FV(M) \subseteq \Gamma \uparrow FV(M \star N)$ and $\Gamma
  \uparrow FV(N) \subseteq \Gamma \uparrow FV(M \star N)$, so $\Gamma \uparrow FV(M \star N) \vdash M :
  \textbf{K}(\alpha \to \beta)$ and $\Gamma \uparrow FV(M \star N) \vdash N : \textbf{K}\alpha$ by weakening.

  Then $\Gamma \uparrow FV(M \star N) \vdash M \star N : \textbf{K}\beta$ by $\textbf{K}_{app}$.

@@

  If $\Gamma, x : \alpha \vdash M : \beta$  and $\Gamma \vdash N : \alpha$, then $\Gamma \vdash (M [x := N]) :
  \beta $

@@

  1) Let $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$ and $\Gamma \vdash N : \alpha$.

  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$, then, by generation, $\Gamma, x : A \vdash
  M : \beta$. So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : \beta $, then $\Gamma \vdash
  {\bf pure \: } (M [x := N]) : \textbf{K}\beta$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
  ({\bf pure \: } M) (M [x := N])$
  by substitution definition, so $\Gamma \vdash ({\bf pure \: } M) (M [x := N]) : \textbf{K}\beta$

  2) Let $\Gamma, x : \gamma \vdash M \star N : \textbf{K}\beta$, and $\Gamma \vdash y : \gamma$.

  So, by generation, $\Gamma, x : \gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma, x : \gamma
  \vdash N : \textbf{K}\alpha$.

  Hence $\Gamma, x : \gamma \vdash (M [x := y]) : \textbf{K}(\alpha \to \beta)$ and
  $\Gamma \vdash (N [x := y]): \textbf{K}\alpha$ by hypothesis.

  So $\Gamma \vdash (M [x := y]) \star (N [x := y]) : \textbf{K}\beta$, or, $\Gamma \vdash (M \star N) ([x
  := y]) : \textbf{K}\beta$.

>>>>

  \begin{defin} Intuitionistic normal modal logic {\bf IK}

    2) $\Box (A \to B) \to (\Box A \to \Box B)$;

    3) Rules: MP and necessiation.

  \begin{defin} Translation from {\bf IK} into {\bf IEL}$^{-}$

    1) $\bot^{\circ} = \bot$;

    2) $p^{\circ} = p$;

    3) $(A \alpha B)^{\circ} = A^{\circ} \alpha B^{\circ}$, $\alpha \in \{ \to, \wedge, \vee \}$.

    4) $(\Box A)^{\circ} = {\bf K} A^{\circ}$
  \end{defin}

  \begin{lemma}
    $ $
    ${\bf IK} \vdash A \Rightarrow {\bf IEL}^{-} \vdash A^{\circ}$
  \end{lemma}

  \begin{proof}
    Straightforward induction on the structure of $M$.
  \end{proof}

  It is clearly that we may prove similar fact for ${\bf IEL}$ by the same way.

@@

  \begin{center}
  \begin{prooftree}
  \AxiomC{$ $}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\Gamma , x : A \vdash x : A$}
  \end{prooftree}
  \end{center}

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  \begin{prooftree}
  \AxiomC{$\Gamma, x : A \vdash M : B$}
  \RightLabel{$\rightarrow_i$}
  \UnaryInfC{$\Gamma \vdash \lambda x. M : A \to B$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{ $\Gamma \vdash x : A$ }
  \AxiomC{ $\Gamma \vdash y : B$ }
  \RightLabel{$\times_i$}
  \BinaryInfC{$\Gamma \vdash \langle x, y \rangle : A \times B$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash x : A$}
    \RightLabel{$\textbf{K}_I$}
    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}A $}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}

  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
  \begin{prooftree}
  \AxiomC{$\Gamma \vdash f : A \to B$}
  \AxiomC{$\Gamma \vdash x : A$}
  \RightLabel{$\rightarrow_e$}
  \BinaryInfC{$\Gamma \vdash fx : B$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
  \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
  \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
    \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
    \RightLabel{${\bf let}_{\textbf{K}}$}
    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: N : {\bf K} B$}
  \end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

  7) $FV({\bf let \:\: pure} \:\: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M_i), \text{where $n = |\vec{M}|$}$.
  \end{defin}

@@

  In $\lambda \textbf{K}$ we have the following computational rules:

@@

$\begin{array}{lll}
&\text{1) } (\lambda x. M) N \rightarrow_{\beta} M [x := N]& \\
&\text{2) } \pi_1 \langle M, N \rangle \rightarrow_{\beta} M& \\
&\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
&\text{4) } {\bf let \: pure \: } \vec{y} = \vec{M_2} {\bf \: in \: } ({\bf let \: pure \: } \vec{x} = \vec{M_1} {\bf \: in \: }  N) \rightarrow_{\beta}& \\
&\quad {\bf let \: pure \: } \vec{x} = ({\bf let \: pure \: } \vec{y} = \vec{M_2} \: {\bf in} \: \vec{M_1}) \: {\bf in} \: {\bf let \: pure \: } \vec{x} = \vec{x} {\bf \: in \:} N&
\end{array}$

@@

  Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.

@@

  If $\Gamma \vdash M : A$, then $FV(M) \subseteq dom(\Gamma)$

@@

  Induction on the derivation of $\Gamma \vdash M : A$.

@@

  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$ implies that $\Gamma \vdash M : A$;

@@

  Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$.

@@

  1) Let $\Gamma \vdash x : A$, such that $\Gamma = \Delta, x : A$ and $\Theta \subseteq \Gamma$.
  Let $\Sigma = \Theta \setminus \Gamma$, or, which is the same, $\Sigma = \Theta \setminus \Delta, x :
  A$, then $\Sigma, \Delta, x : A \vdash x : \alpha$, or, $\Theta \vdash x : A$.

  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$ and $\Gamma \subseteq \Theta$.

  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$, then $\Gamma \vdash M : A$ by generation and, by
  hypothesis, $\Theta \vdash M : A$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}A$ by applying
  $\textbf{K}_I$-rule.

  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ and $\Gamma \subseteq \Delta$.

  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.

  By hypothesis we have $\Delta \vdash \vec{M} : \textbf{K}\vec{A}$. So $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.

@@

  If $\Gamma \vdash M : A$, then $\Gamma \uparrow FV(M) \vdash M : A$, where $\Gamma \uparrow
  FV(M)$ is a subcontext of $\Gamma$, such that $dom(\Gamma \uparrow FV(M)) = dom(\Gamma) \cap FV(M)$.

@@

  Induction by derivation. We consider the base of induction and the case with  ${\bf let}_{\textbf{K}}$. The rest
  cases are
  proven by the same way.

@@

  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : A$, $x \in \mathbb{V}$.

  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : A) \uparrow FV(x) = \{ x
  : A\}$, then $x : A \vdash x : A$ by axiom.

  2) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.

  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.

  By hypothesis $\Gamma \uparrow FV(\vec{M}) \vdash \vec{M} : \textbf{K}\vec{A}$.

  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.

@@

  If $\Gamma, x : A \vdash M : B$  and $\Gamma \vdash N : A$, then $\Gamma \vdash (M [x := N]) :
  B $

@@

  1) Let $\Gamma, x : A \vdash {\bf pure \: } M : \textbf{K}B$ and $\Gamma \vdash N : A$.

  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$.

  By generation, $\Gamma, x : A \vdash M : B$.

  So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : B $.
  Then $\Gamma \vdash {\bf pure \: } (M [x := N]) : \textbf{K}B$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
  ({\bf pure \: } M [x := N])$ by substitution definition.

  So $\Gamma \vdash ({\bf pure \: } M [x := N]) : \textbf{K}B$

  2) .
